---
title: "Untitled"
author: "Josh Browning"
date: "July 14, 2016"
output: html_document
---

First, we load the data and the necessary libraries:

```{r}
library(data.table)
library(bit64)
library(venneuler)
library(ggplot2)
library(maps)
library(mapdata)
# events = fread("data/app_events.csv")
# labels = fread("data/app_labels.csv")
train = fread("data/gender_age_train.csv")
test = fread("data/gender_age_test.csv")
```

Define a venn diagram plotting function:

```{r}
myVenn = function(A, B){
    sizeAnotB = length(unique(A[!A %in% B]))
    sizeBnotA = length(unique(B[!B %in% A]))
    AandB = A[A %in% B]
    if(length(AandB) > 0)
        sizeAandB = length(unique(A[A %in% B]))
    else
        sizeAandB = 0
    v = venneuler(c(A=sizeAnotB, B=sizeBnotA, "A&B"=sizeAandB))
    plot(v)
}
```

Let's first confirm that the device id's between test and train do not overlap at all:

```{r, echo=FALSE}
myVenn(train$device_id, test$device_id)
```

So, there's no overlap, just like we would have expected.

## Brand

What if we look at distributions by phone brand?

```{r}
phone_brand = fread("data/phone_brand_device_model.csv")
phone_brand[, train := device_id %in% train$device_id]
phone_brand[, test := device_id %in% test$device_id]
brand_cnts = phone_brand[, .N, by=phone_brand][order(N, decreasing=TRUE), ]
phone_brand[, phone_brand := factor(phone_brand, levels=brand_cnts$phone_brand)]
ggplot(phone_brand, aes(x=phone_brand, fill=train)) + geom_bar() +
    scale_x_discrete(breaks=brand_cnts$phone_brand, labels=rep("", nrow(brand_cnts)))
```

Most phones fall into one of a few different brands, but there are a total of 131 brands!  We can also look at the proportion of users within each brand that are in test/train (again, the below graph is sorted from highest brand by volume on the left to lowest brand by volume on the right):

```{r}
ggplot(phone_brand, aes(x=phone_brand, fill=train)) + geom_bar(position="fill") +
    scale_x_discrete(breaks=brand_cnts$phone_brand, labels=rep("", nrow(brand_cnts)))
```

Let's confirm all device id's fall into either train or test:

```{r}
phone_brand[, all(xor(train, test))]
```

## Location

```{r}
events = fread("data/events.csv", nrow=1000)
events[, train := events$device_id %in% train$device_id]
events[, test := events$device_id %in% test$device_id]
china_map = fortify(map("china", fill=FALSE, plot=FALSE))
ggplot(events, aes(x=longitude, y=latitude, color=test)) + geom_point() +
    geom_polygon(data=china_map, aes(x=long, y=lat, group=region, color=NULL, fill=NULL))
```