---
title: "Exploratory Analysis"
author: "Josh Browning"
date: "July 14, 2016"
output: html_document
---

## Test vs Train

```{r, echo=FALSE}
suppressPackageStartupMessages({
    library(data.table)
    library(ggplot2)
    library(scales)
    library(gridExtra)
    library(bit64)
})
source("R/plot_proportions.R")
```

```{r}
train = fread("data/gender_age_train.csv")
test = fread("data/gender_age_test.csv")
split_rate = nrow(train) / (nrow(train) + nrow(test))
events = fread("data/events.csv")
events[, train := device_id %in% train$device_id]
```

Let's again look at the uneven split in test vs train by latitude and longitude:

```{r}
to_plot = events[, list(count=.N, proportion=mean(train)), by=round(latitude, 1)]
plot_proportions(to_plot, true_prop=split_rate, log=TRUE)
```

Now longitude:

```{r}
to_plot = events[, list(count=.N, proportion=mean(train)), by=round(longitude, 1)]
plot_proportions(to_plot, true_prop=split_rate, log=TRUE)
```

However, we do see a very even split when looking at event counts:

```{r}
event_cnts = events[, .N, by=c("device_id", "train")]
to_plot = event_cnts[, list(count=.N, proportion=mean(train)), by=N]
plot_proportions(to_plot, true_prop=split_rate, log=TRUE)
```

## Creating Folds

Let's assign folds randomly by device_id:

```{r}
set.seed(123)
fold_cnt = 4
train[, fold := sample(rep(1:4, ceiling(nrow(train)/4)), size=nrow(train))]
train_events = merge(events, train[, list(device_id, fold)], by="device_id")
write.csv(train, file="data/gender_age_train_folds.csv", row.names=FALSE)
```

Now, we can plot the same latitude/longitude graphs as before, but instead look at the proportion in fold i vs the proportion not in fold i:

```{r}
p = list()
for(i in 1:4){
    to_plot = train_events[, list(count=.N, proportion=mean(fold==i)), by=round(latitude, 1)]
    p[[i]] = plot_proportions(to_plot, true_prop=0.25, log=TRUE)
}
grid.arrange(p[[1]], p[[2]], p[[3]], p[[4]])
```

We see similar graphs as we originally saw with the test vs train, and the proportion of groups that fall within the confidence interval is also quite low (and roughly equivalent to what we saw in test vs train).  Longitude is similar:

```{r}
p = list()
for(i in 1:4){
    to_plot = train_events[, list(count=.N, proportion=mean(fold==i)), by=round(longitude, 1)]
    p[[i]] = plot_proportions(to_plot, true_prop=0.25, log=TRUE)
}
grid.arrange(p[[1]], p[[2]], p[[3]], p[[4]])
```

And, not surprisingly, event counts are also similar:

```{r}
p = list()
event_cnts = train_events[, .N, by=c("device_id", "fold")]
for(i in 1:4){
    to_plot = event_cnts[, list(count=.N, proportion=mean(fold==i)), by=N]
    p[[i]] = plot_proportions(to_plot, true_prop=0.25, log=TRUE)
}
grid.arrange(p[[1]], p[[2]], p[[3]], p[[4]])
```

# Split of targets

We can also look at the ratio of observations in each target group (although we could also just do stratified sampling to ensure we have the desirable overlap):

```{r}
for(var in unique(train$group)){
    print(train[, list(count=.N, proportion=mean(group==var)), by=fold])
}
```